Using demographic accounts to enhance official population statistics

### 2021-11-07 article.tex

An example of a minimal demographic account is
\begin{center}
\includegraphics[width=0.4\textwidth]{figures_accounts/fig_account_noage.pdf}
\end{center}
The population contains 16 people at time $t$ and 17 people at time $t+1$; during the interval between $t$ and $t+1$, 8 people are born and 7 die. The change in stock equals inflows minus outflows,
\begin{equation*}
  17 - 16 = 8 - 7,
\end{equation*}
as required.

@@

Publishers of age-structured demographic data typically use some variant of the following scheme,
\begin{center}
\includegraphics[width=0.5\textwidth]{figures_accounts/fig_account_withage2.pdf}
\end{center}
Population stocks stocks and flows are all disaggregated by the same age groups, which, in the case above, are ages 0, 1, and 2+. (The account in question might apply to non-human animals, or use a unit of measurement that is longer than a year.) The final age group includes everone over a certain age. Births are disaggregated by the age of the mother, rather than the age of the child being born, which is always 0.

@@

To do for each age group, need to deaths into ones that occurred to cohort already in age group and cohort entering age group:
\begin{center}
\includegraphics[width=0.5\textwidth]{figures_accounts/fig_account_withage.pdf}
\end{center}

\begin{align}
    6 - (5 + 4) & = -0 + -1 + -2 \quad \text{(Age 2+ at $t+1$)} \\
    5 - 7 & = -1 + -1 \quad \text{(Age 1 at $t+1$)} \\
    6 - (3 + 5) & = -2    
\end{align}

- without age
- with age
   - with cohort
   - without explicit cohort
- show the splitting and combining that occurs   
- age, cohort, and region
   - internal migration is tricky
- age, cohort, and region - fixed attribute (sex? birthplace)

>>>>

An example of a minimal demographic account is
\begin{center}
\includegraphics[width=0.3\textwidth]{figures_accounts/fig_account_noage.pdf}
\end{center}
The population contains 16 people at time $t$ and 17 people at time $t+1$; during the interval between $t$ and $t+1$, 8 people are born and 7 die. The change in stock equals inflows minus outflows,
\begin{equation*}
  17 - 16 = 8 - 7,
\end{equation*}
as required.

@@

Publishers of age-structured demographic data typically use some variant of the following scheme,
\begin{center}
\includegraphics[width=0.45\textwidth]{figures_accounts/fig_account_withage_nolex}
\end{center}
Population stocks stocks and flows are all disaggregated by the same age groups, which, in the case above, are ages 0, 1, and 2+. (The account in question might apply to non-human animals, or use a unit of measurement that is longer than a year.) The final age group includes everone over a certain age. Births are disaggregated by the age of the mother, rather than the age of the child being born, which is always 0.

\begin{equation*}
  \begin{split}
    (8+9+6) - (7+4+5) & = (0+7+0) - (3+1+7) \\
    & \quad \quad + (7+7+5) - (1+5+3)
  \end{split}
\end{equation*}

@@

To do for each age group, need to deaths into ones that occurred to cohort already in age group and cohort entering age group:
\begin{center}
\includegraphics[width=0.5\textwidth]{figures_accounts/fig_account_withage_lex.pdf}
\end{center}

\begin{align}
    8 - (3 + 4) & = -2 + 3 - 1 & \text{(Age 0 at $t+1$)} \\   
    9 - 7 & = -(1 + 1) + (4 + 2) - (0 + 2) & \text{(Age 1 at $t+1$)} \\
    6 - (5 + 4) & = -(0 + 3 + 4) + (5 + 2 + 3) - (3 + 1 + 2)  & \text{(Age 2+ at $t+1$)}
\end{align}
